---
title: "Check MLE of Correlation Matrix"
author: "David Gerard"
date: "January 8, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

Suppose we observed a sample covariance matrix
$$
S = 
\left(
\begin{array}{cc}
c& \rho\\
\rho & 1
\end{array}
\right),
$$
where S is p.d., and we want to estimate a correlation matrix
$$
R = 
\left(
\begin{array}{cc}
1 & a\\
a & 1
\end{array}
\right),
$$
where $a$ is between 0 and 1. We want to maximize the following objective function
$$
-logdet(R) - tr(R^{-1}S).
$$

Under unconstrained maximization of $R$ (so the diagonals need not be 1), we get $\hat{R} = S$.

We want to explore if $a$ is equal to $\rho / \sqrt{c}$. If not, then we are updating the correlation matrix incorrectly in mupdog.

# Analysis

We write out our objective function
```{r}
fn <- function(a, c, rho) {
  -log(1 - a^2) - (c - 2 * a * rho + 1) / (1 - a ^ 2)
}
```

Check my function
```{r}
a <- 0.5
c <- 2
rho <- 0.75

R <- matrix(c(1, a, a, 1), nrow = 2)
S <- matrix(c(c, rho, rho, 1), nrow = 2)
-determinant(R, logarithm = TRUE)$modulus - sum(diag(solve(R) %*% S))
fn(a, c, rho)
```

Maximize fn over $a$ and get something other than $\rho / \sqrt{c}$.

```{r}
a <- 0.5
c <- 2
rho <- 0.75

oout <- optim(par = rho / sqrt(c), fn = fn, 
              method = "Brent", lower = 0, upper = 0.999,
              control = list(fnscale = -1),
              c = c, rho = rho)
oout$par
rho / sqrt(c)
oout$value
fn(rho / sqrt(c), c, rho)
```

If we set $c = 1$, we then get $\hat{a} = \rho / \sqrt{c}$.
```{r}
a <- 0.5
c <- 1
rho <- 0.75

oout <- optim(par = rho / sqrt(c), fn = fn, 
              method = "Brent", lower = 0, upper = 0.999,
              control = list(fnscale = -1),
              c = c, rho = rho)
oout$par
rho / sqrt(c)
oout$value
fn(rho / sqrt(c), c, rho)
```

