---
title: "fitPoly Sims"
author: "David Gerard"
date: "May 1, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

I play around with the results of the fitPoly sims.

# Analysis

```{r, message=FALSE}
library(tidyverse)
dat <- as_data_frame(read.csv("../../../reproduce_genotyping/Output/sims_out/sims_out.csv", row.names = NULL))
```


fitPoly has a lot more outlier datasets:
```{r}
longdat <- dat %>% transmute(updog = uham, Li = bham, fitPoly = fpham,
                             allele_freq = allele_freq, od_param = od_param, 
                             bias_val = bias_val) %>%
  gather(key = "Method", value = "PropCorrect", updog:fitPoly)


pl <- ggplot(data = longdat, mapping = aes(y = PropCorrect, 
                                           x = allele_freq, 
                                           color = Method, 
                                           group = Method)) +
  facet_grid(bias_val ~ od_param) +
  theme_bw() +
  theme(strip.background = element_rect(fill = "white")) +
  xlab("Allele Frequency") +
  ylab("Proportion Correct") +
  geom_point(size = 0.1, alpha = 1/2) +
  geom_smooth(color = "black") +
  guides(colour = guide_legend(override.aes = list(size=1.5))) +
  ggthemes::scale_color_colorblind()
print(pl)
```

fitPoly had a lot of datasets where it wouldn't return anything due to their filtering criteria.

```{r}
dat %>% group_by(od_param, bias_val) %>%
  summarize(num_na = mean(is.na(fpham))) %>%
  ggplot(mapping = aes(x = od_param, 
                       y = bias_val, 
                       fill = num_na,
                       label = num_na)) +
  geom_tile() +
  geom_text() +
  theme_bw() +
  scale_y_continuous(breaks = unique(dat$bias_val)) +
  scale_x_continuous(breaks = unique(dat$od_param)) +
  xlab("Overdispersion") +
  ylab("Bias") +
  guides(fill = guide_legend(title = "Proportion\nMissing")) ->
  pl
print(pl)
```


Updog only does slightly worse on the datasets that fitPoly deems are unrecoverable.
```{r}
dat %>% 
  mutate(fitmiss = is.na(fpham)) %>%
  ggplot(mapping = aes(x = fitmiss, y = uham)) +
  geom_boxplot() +
  facet_grid(bias_val ~ od_param) +
  theme_bw() +
  theme(strip.background = element_rect(fill = "white"))
```


Updog's estimated proportion misclassified is somewhat close to being unbiased. But there is a large variance. And some points that are highly misclassified only have a moderate estimate of being misclassified.
```{r}
dat %>% 
  ggplot(mapping = aes(x = 1 - uham, y = uepm)) +
  geom_point() +
  facet_grid(bias_val ~ od_param) +
  theme_bw() +
  theme(strip.background = element_rect(fill = "white")) +
  xlab("Proportion Misclassified") +
  ylab("Estimated Proportion Misclassified") +
  geom_abline(color = "red", lty = 2) +
  ggtitle("updog") ->
  pl
print(pl)
```

fitPoly's estimates look *much* worse.
```{r}
dat %>% 
  ggplot(mapping = aes(x = 1 - fpham, y = fpepm)) +
  geom_point() +
  facet_grid(bias_val ~ od_param) +
  theme_bw() +
  theme(strip.background = element_rect(fill = "white")) +
  xlab("Proportion Misclassified") +
  ylab("Estimated Proportion Misclassified") +
  geom_abline(color = "red", lty = 2) +
  ggtitle("fitPoly") ->
  pl
print(pl)
```


fitPoly has some poor behavior in the correlation metric.
```{r}
longdat <- dat %>% transmute(updog = ucor, Li = bcor, fitPoly = fpcor,
                             allele_freq = allele_freq, od_param = od_param, 
                             bias_val = bias_val) %>%
  gather(key = "Method", value = "Cor", updog, Li, fitPoly)


pl <- ggplot(data = longdat, mapping = aes(y = Cor, 
                                           x = allele_freq, 
                                           color = Method, 
                                           group = Method)) +
  facet_grid(bias_val ~ od_param) +
  theme_bw() +
  theme(strip.background = element_rect(fill = "white")) +
  xlab("Allele Frequency") +
  ylab("Correlation with True Genotype") +
  geom_point(size = 0.1, alpha = 1/2) +
  geom_smooth(color = "black") +
  guides(colour = guide_legend(override.aes = list(size=1.5))) +
  ggthemes::scale_color_colorblind()
print(pl)
```

```{r}
longdat %>%
  ggplot(mapping = aes(x = Method, y = Cor)) +
  facet_grid(bias_val ~ od_param) +
  theme_bw() +
  theme(strip.background = element_rect(fill = "white")) +
  ylab("Correlation with True Genotype") +
  geom_boxplot() ->
  pl
print(pl)
```

