---
title: "updog vs flexdog"
author: "David Gerard"
date: "March 13, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

This is to prove that `updog` with option `model = "hw"` is the same as `flexdog` with option `model = "hw"`.

# Analysis
```{r, message=FALSE}
library(updog)
library(mupdog)
library(tidyverse)
```

## SNP 1
```{r}
data("snpdat")
ploidy  <- 6
snpdat %>% 
  filter(snp == "SNP1") ->
  smalldat
refvec  <- smalldat$counts
refvec  <- refvec[!is.na(refvec)]
sizevec <- smalldat$size
sizevec <- sizevec[!is.na(sizevec)]
```

Fit both `updog` and `flexdog` with the HW options.
```{r}
uout <- updog_vanilla(ocounts        = refvec, 
                      osize          = sizevec, 
                      ploidy         = ploidy, 
                      update_outprop = FALSE, 
                      out_prop       = 0, 
                      model          = "hw", 
                      seq_error_mean = -4.7, 
                      seq_error_sd   = 1, 
                      bias_val_mean  = 0, 
                      bias_val_sd    = 0.7)

fout <- flexdog(refvec    = refvec, 
                sizevec   = sizevec, 
                ploidy    = ploidy, 
                model     = "hw",
                verbose   = FALSE, 
                mean_bias = 0,
                var_bias  = 0.7^2, 
                mean_seq  = -4.7, 
                var_seq   = 1,
                tol       = 10^-5)
```

Estimates are nearly exactly the same!
```{r}
all(fout$geno == uout$ogeno)
fout$bias_val
uout$bias_val

fout$od_param
uout$od_param

fout$seq_error
uout$seq_error

fout$par$alpha
uout$allele_freq
```

## SNP 2

```{r}
data("snpdat")
ploidy  <- 6
snpdat %>% 
  filter(snp == "SNP2") ->
  smalldat
refvec  <- smalldat$counts
refvec  <- refvec[!is.na(refvec)]
sizevec <- smalldat$size
sizevec <- sizevec[!is.na(sizevec)]
```

Fit both `updog` and `flexdog` with the HW options.
```{r}
uout <- updog_vanilla(ocounts        = refvec, 
                      osize          = sizevec, 
                      ploidy         = ploidy, 
                      update_outprop = FALSE, 
                      out_prop       = 0, 
                      model          = "hw", 
                      seq_error_mean = -4.7, 
                      seq_error_sd   = 1, 
                      bias_val_mean  = 0, 
                      bias_val_sd    = 0.7)

fout <- flexdog(refvec    = refvec, 
                sizevec   = sizevec, 
                ploidy    = ploidy, 
                model     = "hw",
                verbose   = FALSE, 
                mean_bias = 0,
                var_bias  = 0.7^2, 
                mean_seq  = -4.7, 
                var_seq   = 1,
                tol       = 10^-5)
```

Estimates are nearly exactly the same!
```{r}
all(fout$geno == uout$ogeno)
fout$bias_val
uout$bias_val

fout$od_param
uout$od_param

fout$seq_error
uout$seq_error

fout$par$alpha
uout$allele_freq
```

## SNP 3

```{r}
data("snpdat")
ploidy  <- 6
snpdat %>% 
  filter(snp == "SNP3") ->
  smalldat
refvec  <- smalldat$counts
refvec  <- refvec[!is.na(refvec)]
sizevec <- smalldat$size
sizevec <- sizevec[!is.na(sizevec)]
```

Fit both `updog` and `flexdog` with the HW options.
```{r}
uout <- updog_vanilla(ocounts        = refvec, 
                      osize          = sizevec, 
                      ploidy         = ploidy, 
                      update_outprop = FALSE, 
                      out_prop       = 0, 
                      model          = "hw", 
                      seq_error_mean = -4.7, 
                      seq_error_sd   = 1, 
                      bias_val_mean  = 0, 
                      bias_val_sd    = 0.7)

fout <- flexdog(refvec    = refvec, 
                sizevec   = sizevec, 
                ploidy    = ploidy, 
                model     = "hw",
                verbose   = FALSE, 
                mean_bias = 0,
                var_bias  = 0.7^2, 
                mean_seq  = -4.7, 
                var_seq   = 1,
                tol       = 10^-5)
```

Estimates are nearly exactly the same!
```{r}
all(fout$geno == uout$ogeno)
fout$bias_val
uout$bias_val

fout$od_param
uout$od_param

fout$seq_error
uout$seq_error

fout$par$alpha
uout$allele_freq
```

