---
title: "FIrst Try at Flexdog"
author: "David Gerard"
date: "March 13, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

Here, I try out `flexdog` for the first time. I then compare the fits to that of `mupdog`.

# Analysis

First, I read in the mupdog fit.

```{r, message=FALSE}
library(mupdog)
library(tidyverse)
mout <- readRDS("../output/shir_fit/shir_fit_nodiag1.RDS")
```

Now I fit `flexdog` on the first few SNPs.

## SNP 1

Extract SNP1 data.
```{r, cache=TRUE}
refvec  <- mout$input$refmat[, 1]
sizevec <- mout$input$sizemat[, 1]
ploidy  <- mout$input$ploidy
plot(mout, 1) +
  ggtitle("mupdog fit")
aout <- flexdog(refvec = refvec, sizevec = sizevec, 
                ploidy = ploidy, model = "ash")
plot(aout)  +
    ggtitle("flexdog fit")
plot(aout$gene_dist, type = "h", main = "Genotype Distribution", lwd = 4)
```



## SNP 2
```{r, cache=TRUE}
refvec  <- mout$input$refmat[, 2]
sizevec <- mout$input$sizemat[, 2]
refvec  <- refvec[!is.na(refvec)]
sizevec <- sizevec[!is.na(sizevec)]
plot(mout, 2) +
  ggtitle("mupdog fit")
aout <- flexdog(refvec = refvec, sizevec = sizevec, 
                ploidy = ploidy, model = "ash")
plot(aout) +
    ggtitle("flexdog fit")
plot(aout$gene_dist, type = "h", main = "Genotype Distribution", lwd = 4)
```


The estimate for the bias is very unstable. We probably need more regularization on the bias parameter.

```{r, cache = TRUE}
aout <- flexdog(refvec = refvec, sizevec = sizevec, 
                ploidy = ploidy, model = "ash",
                var_bias = 0.01)
plot(aout) +
  ggtitle("flexdog fit var_bias = 0.01")
plot(aout$gene_dist, type = "h", main = "Genotype Distribution", lwd = 4)
```

So, this tells me for `flexdog`, we should estimate `var_bias` adaptively, rather than setting it.

## SNP 3
```{r, cache=TRUE}
refvec  <- mout$input$refmat[, 3]
sizevec <- mout$input$sizemat[, 3]
refvec  <- refvec[!is.na(refvec)]
sizevec <- sizevec[!is.na(sizevec)]
plot(mout, 3) +
  ggtitle("mupdog fit")
aout <- flexdog(refvec = refvec, sizevec = sizevec, 
                ploidy = ploidy, model = "ash")
plot(aout) +
    ggtitle("flexdog fit")
plot(aout$gene_dist, type = "h", main = "Genotype Distribution", lwd = 4)
```

Again, let's increase the regularization on the bias.
```{r, cache = TRUE}
aout <- flexdog(refvec = refvec, sizevec = sizevec, 
                ploidy = ploidy, model = "ash",
                var_bias = 0.01)
plot(aout) +
  ggtitle("flexdog fit var_bias = 0.01")
plot(aout$gene_dist, type = "h", main = "Genotype Distribution", lwd = 4)
```

## SNP 4
```{r, cache=TRUE}
refvec  <- mout$input$refmat[, 4]
sizevec <- mout$input$sizemat[, 4]
refvec  <- refvec[!is.na(refvec)]
sizevec <- sizevec[!is.na(sizevec)]
plot(mout, 4) +
  ggtitle("mupdog fit")
aout <- flexdog(refvec = refvec, sizevec = sizevec, 
                ploidy = ploidy, model = "ash")
plot(aout) +
    ggtitle("flexdog fit")
plot(aout$gene_dist, type = "h", main = "Genotype Distribution", lwd = 4)
```


## SNP 5
```{r, cache=TRUE}
refvec  <- mout$input$refmat[, 5]
sizevec <- mout$input$sizemat[, 5]
refvec  <- refvec[!is.na(refvec)]
sizevec <- sizevec[!is.na(sizevec)]
plot(mout, 5) +
  ggtitle("mupdog fit")
aout <- flexdog(refvec = refvec, sizevec = sizevec, 
                ploidy = ploidy, model = "ash")
plot(aout) +
    ggtitle("flexdog fit")
plot(aout$gene_dist, type = "h", main = "Genotype Distribution", lwd = 4)
```




## High Bias SNP
```{r, cache=TRUE}
refvec  <- mout$input$refmat[, 152]
sizevec <- mout$input$sizemat[, 152]
refvec  <- refvec[!is.na(refvec)]
sizevec <- sizevec[!is.na(sizevec)]
plot(mout, 152) +
  ggtitle("mupdog fit")
aout <- flexdog(refvec = refvec, sizevec = sizevec, 
                ploidy = ploidy, model = "ash")
plot(aout) +
    ggtitle("flexdog fit")
plot(aout$gene_dist, type = "h", main = "Genotype Distribution", lwd = 4)
```
