---
title: "How Different Does PP Make the Prior"
author: "David Gerard"
date: "May 29, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

We developed a model for preferential pairing. How different does this make the genotype distribution compared to the absence of preferential pairing?

# Analysis

We will assume we are in an S1 population tetraploids. In this case, the only parental genotype that results in a different inheritance distribution is 2. Let's look at how extreme the difference can be.
```{r, message=FALSE}
library(updog)
library(tidyverse)
library(ggthemes)
ploidy <- 4
p1geno <- p2geno <- 2
blist <- get_bivalent_probs(ploidy = ploidy)
p1dist <- blist$probmat[blist$lvec == p1geno, , drop = FALSE]
p2dist <- blist$probmat[blist$lvec == p2geno, , drop = FALSE]
```

These are the weights needed to get back non-preferential pairing:
```{r}
get_hyper_weights(ploidy = ploidy, ell = p1geno)$weightvec
```

Let's put all of the weight on the first or second second term
```{r}
weightvec  <- c(0, 1)
p1segprob  <- colSums(weightvec * p1dist)
p2segprob  <- colSums(weightvec * p2dist)
gene_dist1 <- c(0, 0, 1, 0, 0) ## all weight on first
gene_dist2 <- c(updog::convolve(x = p1segprob, y = p2segprob))
f1dist     <- get_q_array(ploidy = ploidy)[p1geno + 1, p2geno + 1, ]
```

```{r}
data_frame(None            = f1dist,
           "All on First"  = gene_dist1,
           "All on Second" = gene_dist2, 
           geno            = 0:4) %>%
  gather(key = "Type", value = "Probability", 1:3) ->
  smalldat
smalldat$geno[smalldat$Type == "None"] <-
  smalldat$geno[smalldat$Type == "None"] - 0.1
smalldat$geno[smalldat$Type == "All on Second"] <-
  smalldat$geno[smalldat$Type == "All on Second"] + 0.1

ggplot(data = smalldat,
  mapping = aes(x     = geno, 
                xend  = geno, 
                y     = 0, 
                yend  = Probability,
                lty   = Type,
                color = Type)) +
  geom_segment(lineend = "round", lwd = 1) +
  theme_bw() +
  xlab("Genotype") +
  ylab("Probability") +
  ggthemes::scale_color_colorblind()
```


## Aside: comparing convolutions

My convolution is way faster for this scale:
```{r}
microbenchmark::microbenchmark(
stats::convolve(p1segprob, rev(p2segprob), type = "o"),
updog::convolve(x = p1segprob, y = p2segprob)
)
```




