---
title: "Hypergeometric as Special Case"
author: "David Gerard"
date: "May 23, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

I demonstrate that the hypergeometric distribution is a special case of the mixture of off-zero binomials.

# Demonstration

```{r}
library(updog)
ploidy <- 10
ell    <- 4 # allele dosage

hdist <- stats::dhyper(x = 0:(ploidy / 2),
                       m = ell,
                       n = ploidy - ell, 
                       k = ploidy / 2)
blist <- get_bivalent_probs(ploidy = ploidy)
hweight_list <- get_hyper_weights(ploidy = ploidy, ell = ell)


distmat <- blist$probmat[blist$lvec == ell, , drop = FALSE]
distmat
weightvec <- hweight_list$weightvec
weightvec
```

The hypergeometric distribution and the mixture distribution are the exact same:
```{r}
colSums(distmat * weightvec)
hdist
```

Thus, if there were no preferential pairing, we would expect the following configurations
```{r}
hweight_list$pmat
```
with probabilities:
```{r}
hweight_list$weightvec
```

Deviations away from these mixing probabilities indicate preferential pairing.


