---
title: "Check Gaussian Binomial Dist"
author: "David Gerard"
date: "March 8, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

In `mupdog` we are using a Gaussian-Binomial distribution to estimate the distribution of genotypes. 
\begin{align*}
z &\sim N(0, \sigma^2)\\
y &= F^{-1}(\Phi(z)|K, \alpha),
\end{align*}
where $F^{-1}(\cdot|K,\alpha)$ is the quantile function of a binomial distribution with size $K$ and success probability $\alpha$, and $\Phi(\cdot)$ is the CDF of a standard normal.

Let's see how good this model can possibly be under model mispecification.

# Analysis

Since we are estimating parameters by maximum likelihood, as we get more individuals the estimate of the Gaussian-Binomial distribution will converge to that which is closes to the true distribution in terms of Kullback-Leibler divergence. I made the function `kl_opt` in `mupdog` to find this minimizer for any provided distribution `dist`.

Let's suppose that the individuals are actually all siblings, let's compare the resulting distribution across a range of genotypes.
```{r}
ploidy <- 6
qarray <- updog::get_q_array(ploidy)
```

## Parental genotypes = 1 and 2
Let's say the parents have genotype Aaaaaa and AAaaaaa
```{r}
dist <- qarray[2, 3, ]
klout <- mupdog::kl_opt(dist = dist, ploidy = ploidy)
```

The optimal values are
```{r}
klout$sigma2
klout$alpha
```

The value of $\alpha$ is close to the mean of `r sum(dist * (0:ploidy)) / ploidy`.

The distributions looks pretty close! The binomial looks a lot worse.
```{r, message=FALSE}
library(tidyverse)
library(ggthemes)
bindist <- dbinom(x = 0:ploidy, size = ploidy, prob = sum(dist * (0:ploidy)) / ploidy)
data_frame(dosage = 0:ploidy, true = dist, approximation = klout$dist,
           binomial = bindist) %>%
  gather(key = "distribution", value = "probability", 2:4) ->
  dfdat
dfdat$dosage[dfdat$distribution == "approximation"] <- dfdat$dosage[dfdat$distribution == "approximation"] + 0.1
dfdat$dosage[dfdat$distribution == "binomial"] <- dfdat$dosage[dfdat$distribution == "binomial"] + 0.2
ggplot(dfdat, mapping = aes(x = dosage, xend = dosage, y = 0, yend = probability, lty = distribution,
                            color = distribution)) +
  geom_segment() +
  theme_bw() +
  ylab("probability") +
  scale_color_colorblind()
```

## Parents of a 2, 2

```{r}
dist <- qarray[3, 3, ]
klout <- mupdog::kl_opt(dist = dist, ploidy = ploidy)
```

The optimal values are
```{r}
klout$sigma2
klout$alpha
```

The value of $\alpha$ is close to the mean of `r sum(dist * (0:ploidy)) / ploidy`.

The distributions looks pretty close! The binomial looks a lot worse.
```{r, message=FALSE}
library(tidyverse)
library(ggthemes)
bindist <- dbinom(x = 0:ploidy, size = ploidy, prob = sum(dist * (0:ploidy)) / ploidy)
data_frame(dosage = 0:ploidy, true = dist, approximation = klout$dist,
           binomial = bindist) %>%
  gather(key = "distribution", value = "probability", 2:4) ->
  dfdat
dfdat$dosage[dfdat$distribution == "approximation"] <- dfdat$dosage[dfdat$distribution == "approximation"] + 0.1
dfdat$dosage[dfdat$distribution == "binomial"] <- dfdat$dosage[dfdat$distribution == "binomial"] + 0.2
ggplot(dfdat, mapping = aes(x = dosage, xend = dosage, y = 0, yend = probability, lty = distribution,
                            color = distribution)) +
  geom_segment() +
  theme_bw() +
  ylab("probability") +
  scale_color_colorblind()
```

## Iterate through all possible parents

```{r}
sim_mat <- matrix(NA, nrow = (ploidy + 1) * (ploidy + 2) / 2, ncol = 5)
true_list <- list()
approx_list <- list()
index <- 1
for (i in 0:ploidy) {
  for (j in i :ploidy) {
    dist <- qarray[i + 1, j + 1, ]
    klout <- mupdog::kl_opt(dist = dist, ploidy = ploidy)
    sim_mat[index, 1] <- i
    sim_mat[index, 2] <- j
    sim_mat[index, 3] <- klout$sigma2
    sim_mat[index, 4] <- klout$alpha
    sim_mat[index, 5] <- sum((0:ploidy) * dist / ploidy)
    true_list[[index]] <- dist
    approx_list[[index]] <- klout$dist
    
    dist_df <- data_frame(dosage = 0:ploidy, true = dist, approximation = klout$dist) %>%
      gather(key = "distribution", value = "probability", 2:3)
      dist_df$dosage[dist_df$distribution == "approximation"] <- dist_df$dosage[dist_df$distribution == "approximation"] + 0.2
      dist_df$ij <- paste0("(", i, ",", j, ")")
    
    if (i == 0 & j == 0) {
      dist_tot <- dist_df
    } else {
      dist_tot <- bind_rows(dist_tot, dist_df)
    }
    index <- index + 1
  }
}
sim_df <- as_data_frame(sim_mat)
names(sim_df) <- c("i", "j", "sigma2", "alpha", "true_mean")
```

```{r, fig.height=30}
ggplot(dist_tot, mapping = aes(x = dosage, xend = dosage, y = 0, yend = probability, lty = distribution,
                            color = distribution)) +
  geom_segment() +
  theme_bw() +
  ylab("probability") +
  scale_color_colorblind() +
  facet_wrap(~ij, ncol = 2) +
  theme(strip.background = element_rect(fill = "white"))
```

# Different Sigma

However, we generally only have one sigma for each individual shared across all SNPs. So we can't have it be so flexible.

```{r}
hist(sim_df$sigma2)
mean(sim_df$sigma2)
```

Let's see how they all look when we fix $\sigma^2$ to be 0.4 and we let $\alpha$ be the mean of the true distribution.

It looks OK, but there are some combinations that don't look so good.

```{r}
sigma2 <- 0.4
for (i in 0:ploidy) {
  for (j in i:ploidy) {
    dist <- qarray[i + 1, j + 1, ]
    alpha <- sum(dist * (0:ploidy)) / ploidy
    approx <- c(mupdog:::compute_all_post_prob(ploidy = ploidy, 
                                      mu = matrix(data = 0, nrow = 1, ncol = 1), 
                                      sigma2 = matrix(data = sigma2, nrow = 1, ncol = 1), 
                                      alpha = alpha, rho = 0))
    
    dist_df <- data_frame(dosage = 0:ploidy, true = dist, approximation = approx) %>%
      gather(key = "distribution", value = "probability", 2:3)
      dist_df$dosage[dist_df$distribution == "approximation"] <- dist_df$dosage[dist_df$distribution == "approximation"] + 0.2
      dist_df$ij <- paste0("(", i, ",", j, ")")
      dist_df$alpha <- alpha
      dist_df$sigma2 <- sigma2
      
    if (i == 0 & j == 0) {
      dist_tot <- dist_df
    } else {
      dist_tot <- bind_rows(dist_tot, dist_df)
    }
  }
}
```


```{r, fig.height=30}
ggplot(dist_tot, mapping = aes(x = dosage, xend = dosage, y = 0, yend = probability, lty = distribution,
                            color = distribution)) +
  geom_segment() +
  theme_bw() +
  ylab("probability") +
  scale_color_colorblind() +
  facet_wrap(~ij, ncol = 2) +
  theme(strip.background = element_rect(fill = "white"))
```

