---
title: "Normal BB"
author: "David Gerard"
date: "January 9, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

I explore properties of what I will call the normal-beta-binomial model. Specifically, this is a generative model defined by the two-step process:
$$
z \sim N(0, \sigma^2),\\
y = F^{-1}(\Phi(z)|\alpha, \tau),
$$
where $F^{-1}(\cdot | \alpha, \tau)$ is the quantile function of a beta-binomial random variable with mean $\alpha$ and overdispersion parameter $\tau$, and $\Phi(\cdot)$ is the *standard* normal CDF.

# Analysis

First, make sure that the normal copula model does indeed result in bb
```{r}
library(mupdog)
library(tidyverse)
K     <- 6
mu    <- 5/6
rho   <- 0.01
n     <- 400000
sigma <- 1

z <- rnorm(n = n, mean = 0, sd = sigma)
simout_nbb <- qbetabinom(p = pnorm(z), size = K, mu = mu, rho = rho)

emp_dist <- c(table(simout_nbb) / n)
theo_dist <- dbetabinom(0:K, K, mu, rho, FALSE)

qplot(emp_dist, theo_dist) + theme_bw() + geom_abline(lty = 2)

mean(simout_nbb) ## should be about 5
```


What if we vary the SD. We don't still get $K\mu$.
```{r, cache=TRUE}
K <- 6
mu <- 5/6
rho <- 0.01
n <- 400000
sigmaseq <- 10 ^ seq(-2, 3, length = 20)
xbar <- rep(NA, length(sigmaseq))
for (index in 1:length(xbar)) {
  z <- rnorm(n = n, mean = 0, sd = sigmaseq[index])
  simout_nbb <- qbetabinom(p = pnorm(z), size = K, mu = mu, rho = rho)
  xbar[index] <- mean(simout_nbb)
}

qplot(log(sigmaseq, base = 10), xbar, xlab = "log (base 10) sd", ylab = "mean", geom = "line", main = "Mean of BB is 5") + theme_bw()
```

Interestingly, when you decrease the variance a little bit the mean goes up then goes back down to 5. When you increase the variance, the mean decreases toward $K/2 = 3$, which makes intuitive sense.

What about a different value of $\mu$? Say, $\mu = 1/6$?

```{r, cache=TRUE}
K <- 6
mu <- 1/6
rho <- 0.01
n <- 400000
sigmaseq <- 10 ^ seq(-2, 3, length = 20)
xbar <- rep(NA, length(sigmaseq))
for (index in 1:length(xbar)) {
  z <- rnorm(n = n, mean = 0, sd = sigmaseq[index])
  simout_nbb <- qbetabinom(p = pnorm(z), size = K, mu = mu, rho = rho)
  xbar[index] <- mean(simout_nbb)
}

qplot(log(sigmaseq, base = 10), xbar, xlab = "log (base 10) sd", ylab = "mean", geom = "line", main = "Mean of BB is 5") + theme_bw()
```

We see a mirror image.

At $\mu = 1/2$, the mean stays around 3 (notice the scales on the Y-axis).

```{r, cache=TRUE}
K <- 6
mu <- 3/6
rho <- 0.01
n <- 400000
sigmaseq <- 10 ^ seq(-2, 3, length = 20)
xbar <- rep(NA, length(sigmaseq))
for (index in 1:length(xbar)) {
  z <- rnorm(n = n, mean = 0, sd = sigmaseq[index])
  simout_nbb <- qbetabinom(p = pnorm(z), size = K, mu = mu, rho = rho)
  xbar[index] <- mean(simout_nbb)
}

qplot(log(sigmaseq, base = 10), xbar, xlab = "log (base 10) sd", ylab = "mean", geom = "line", main = "Mean of BB is 5") + theme_bw()
```

```{r}
sessionInfo()
```

