---
title: "Outlier Comparision between Updog and Mupdog"
author: "David Gerard"
date: "May 1, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

I compare `flexdog` and `updog` in terms of their outlier estimates. The prob_outlier is mostly the same.

But other summaries are very different (see SNP 3)! 
This is because flexdog says that you have no genotype if you are an outlier while updog says you have a genotype, but it isn't observed. Very subtle!

That is, `updog` says that given genotype k, you are either BB with mean at k/K or BB from the outlier distribution.

While `flexdog` says that you have genotype k with some probabilitiy or you have no genotype with some probability, and thus follow the outlier distribution.

`updog` seems more philosophically sound, but the genotyping results seem similar.

# Analysis
```{r, message=FALSE}
library(updog)
library(mupdog)
library(tidyverse)
data("snpdat")
```

# SNP 1
```{r}
snpdat %>%
  filter(snp == "SNP1") ->
  smalldat
refvec <- smalldat$counts[-1]
sizevec <- smalldat$size[-1]
p1ref <- smalldat$counts[1]
p1size <- smalldat$size[1]
ploidy <- 6
fout <- flexdog(refvec   = refvec, 
                sizevec  = sizevec,
                ploidy   = ploidy,
                model    = "s1", 
                p1ref    = p1ref,
                p1size   = p1size,
                outliers = TRUE)
uout <- updog(ocounts  = refvec, 
              osize    = sizevec, 
              ploidy   = ploidy, 
              model    = "s1", 
              p1counts = p1ref, 
              p1size   = p1size)
```

```{r}
fout$out_prop
uout$out_prop

qplot(fout$prob_outlier, uout$prob_out) +
  geom_abline() +
  ggtitle("prob outlier")

qplot(uout$maxpostprob, fout$maxpostprob) +
  geom_abline() +
  ggtitle("maxpostprob")

qplot(uout$postmean, fout$postmean) +
  geom_abline() +
  ggtitle("Post mean")

plot(fout)
plot(uout, use_colorblind = TRUE)
```

# SNP 2

```{r}
snpdat %>%
  filter(snp == "SNP2") ->
  smalldat
refvec <- smalldat$counts[-1]
sizevec <- smalldat$size[-1]
p1ref <- smalldat$counts[1]
p1size <- smalldat$size[1]
ploidy <- 6
fout <- flexdog(refvec   = refvec, 
                sizevec  = sizevec,
                ploidy   = ploidy,
                model    = "s1", 
                p1ref    = p1ref,
                p1size   = p1size,
                outliers = TRUE)
uout <- updog(ocounts  = refvec, 
              osize    = sizevec, 
              ploidy   = ploidy, 
              model    = "s1", 
              p1counts = p1ref, 
              p1size   = p1size)
```

```{r}
fout$out_prop
uout$out_prop

qplot(fout$prob_outlier, uout$prob_out) +
  geom_abline() +
  ggtitle("prob outlier")

qplot(uout$maxpostprob, fout$maxpostprob) +
  geom_abline() +
  ggtitle("maxpostprob")

qplot(uout$postmean, fout$postmean) +
  geom_abline() +
  ggtitle("Post mean")

plot(fout)
plot(uout, use_colorblind = TRUE)
```


# SNP 3

```{r}
snpdat %>%
  filter(snp == "SNP3") ->
  smalldat
refvec <- smalldat$counts[-1]
sizevec <- smalldat$size[-1]
p1ref <- smalldat$counts[1]
p1size <- smalldat$size[1]
ploidy <- 6
fout <- flexdog(refvec   = refvec, 
                sizevec  = sizevec,
                ploidy   = ploidy,
                model    = "s1", 
                p1ref    = p1ref,
                p1size   = p1size,
                outliers = TRUE)
uout <- updog(ocounts  = refvec, 
              osize    = sizevec, 
              ploidy   = ploidy, 
              model    = "s1", 
              p1counts = p1ref, 
              p1size   = p1size)
```

```{r}
fout$out_prop
uout$out_prop

qplot(fout$prob_outlier, uout$prob_out) +
  geom_abline() +
  ggtitle("prob outlier")

qplot(uout$maxpostprob, fout$maxpostprob) +
  geom_abline() +
  ggtitle("maxpostprob")

qplot(uout$postmean, fout$postmean) +
  geom_abline() +
  ggtitle("Post mean")

plot(fout)
plot(uout, use_colorblind = TRUE)
```




```{r}
sweep(x = fout$postmat, MARGIN = 1, STATS = rowSums(fout$postmat), FUN = "/") ->
  fpostmat

qplot(c(fpostmat), c(uout$postmat)) +
  geom_abline()


```

