---
title: "Explore Shirasawa Correlation More Carefully"
author: "David Gerard"
date: "January 10, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

I take a look more closely at the Shirasawa correlation matrix.

# Analysis

```{r, message=FALSE, warning=FALSE}
library(corrplot)
library(tidyverse)
mout <- readRDS("../output/shir_fit/shir_fit_nodiag1.RDS")
R <- cov2cor(mout$cor_mat)
```


```{r, echo = FALSE, eval = FALSE}
# The correlations between the parent and the offspring are different 
# than the correlations between the offspring and the offspring. Though not by much (and can't believe that df/p-value).
par <- R[1, -1]
nonpar <- R[-1, -1][upper.tri(R[-1, -1])]
t.test(par, nonpar)
boxplot(par, nonpar, names = c("Parent", "No Parent"))
```

The genotype correlation looks as we would expect, with higher correlation of the parent with everyone else:
```{r}
genocor <- cor(t(mout$map_dosage))
corrplot(genocor, method = "color", order = "hclust", diag = FALSE)
qplot(genocor[upper.tri(genocor)], bins = 30) + theme_bw()
data_frame(cor = genocor[1, -1], parent = TRUE) %>%
  bind_rows(data_frame(cor = genocor[-1, -1][upper.tri(genocor[-1, -1])], parent = FALSE)) %>%
  ggplot(mapping = aes(x = parent, y = cor)) +
  geom_boxplot() +
  theme_bw()
```

But this is not what `R` is estimating. Rather, we also must multiply the genotypes on the left by the centering matrix prior to calculating the correlation.
```{r}
n <- nrow(genocor)
L <- diag(n) - matrix(1 / n, nrow = n, ncol = n)
cordiff <- cor(t(L %*% mout$map_dosage))
corrplot(cordiff, method = "color", order = "hclust", diag = FALSE)
```

What is the concordance here between `cordiff` and `R`?

```{r}
plot(R[upper.tri(R)], cordiff[upper.tri(cordiff)])
```

