---
title: "Try mupdog on top 1000 SNPs"
author: "David Gerard"
date: "December 20, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

I try out an inefficient version of mupdog for the first time.

# Analysis

```{r}
library(tidyverse)
library(mupdog)
refmat <- as.matrix(read.csv("/home/david/Code/reproduce_genotyping/Output/shirasawa_snps/example_refcounts.csv", row.names = 1))
sizemat <- as.matrix(read.csv("/home/david/Code/reproduce_genotyping/Output/shirasawa_snps/example_readcounts.csv", row.names = 1))
```

Subset since I think this will be slow

```{r}
nind <- nrow(refmat)
nsnps <- ncol(refmat)
```

Try it out!
```{r, cache = TRUE}
mout <- mupdog::mupdog(refmat = refmat, sizemat = sizemat, ploidy = 6, verbose = FALSE, 
                       control = list(obj_tol = 10^-5), num_clust = 7)
```

# Look at genotyping

The correlation matrix looks a little different than what I expected --- I expected all of the correlations to be about the same.
```{r}
hist(mout$cor_mat[upper.tri(mout$cor_mat)])
```

```{r}
hist(mout$seq)
```

```{r}
hist(log(mout$bias))
```

```{r}
hist(mout$od)
```

Inbreeding doesn't seem to be too important for these data
```{r}
hist(log(mout$inbreeding))
```

```{r}
## qwraps2::lazyload_cache_dir(path = "./try_mupdog_1000_cache/html/")
library(corrplot)
cororder <- corrMatOrder(mout$cor_mat, order = "hclust")
cormat <- mout$cor_mat[cororder, cororder]
corrplot(mout$cor_mat, method = "color", diag = FALSE, order = "hclust", tl.pos = "n")
```


The Genotypes all look fantastic.
```{r, cache=TRUE}
for (index in 1:10) {
  pl <- plot(mout, index)
  print(pl)
}

for (index in 991:1000) {
  pl <- plot(mout, index)
  print(pl)
}
```





```{r}
sessionInfo()
```















