---
title: "Test BB option on Shirasawa Data"
author: "David Gerard"
date: "April 18, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

I test the new BB option for the genotype distribution in the Shirasawa data.

The OD parameter is estimated to be near the minimum I allow. This makes sense since the Shirasawa data consists of siblings, which have *less* dispersed genotypes than a binomial, not more like the assumption when using the "bb" option.

# Analysis

```{r}
library(mupdog)
library(updog)
library(tidyverse)
data("snpdat")
```

SNP 1
```{r}
smalldat <- filter(snpdat, snp == "SNP1")
fout <- flexdog(refvec = smalldat$counts, sizevec = smalldat$size, ploidy = 6, model = "bb")
fout$par
```

SNP 2
```{r}
smalldat <- filter(snpdat, snp == "SNP2")
fout <- flexdog(refvec = smalldat$counts, sizevec = smalldat$size, ploidy = 6, model = "bb")
fout$par
```

SNP 3
```{r}
smalldat <- filter(snpdat, snp == "SNP3")
fout <- flexdog(refvec = smalldat$counts, sizevec = smalldat$size, ploidy = 6, model = "bb")
fout$par
```


# Demonstrate that HW is more dispersed
```{r}
true_dist <- updog::get_q_array(6)[6,6, ]
hwe_dist <- dbinom(0:6, 6, 5/6)
plot(0:6, true_dist, type = "h")
points(0:6 + 0.1, hwe_dist, type = "h", col = 2, lty = 2)
legend("topleft", lty = c(1, 2), col = c(1, 2), legend = c("S1", "HWE"))
```

