---
title: "uit_nodiag"
author: "David Gerard"
date: "January 10, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

I fit the uitdewilligen dataset to the new version of mupdog that does not constrain the diagonal of R to be 1.

See [here](uitdewilligen.html) for fitting while constraining the diagonal to be 1.

# Analysis

```{r, message=FALSE, warning=FALSE}
library(mupdog)
library(tidyverse)
library(corrplot)
```


```{r, eval = FALSE}
mnocor <- readRDS("../output/uit_fit/fit_uit1.RDS")
mout <- mupdog(refmat = mnocor$input$refmat, sizemat = mnocor$input$sizemat, ploidy = mnocor$input$ploidy, 
               verbose = TRUE, num_core = 7,
               control = list(obj_tol = 10^-5, itermax = 200))
saveRDS(mout, "../output/uit_fit/fit_uit_nodiag.RDS")
```


```{r}
mout_nodiag <- readRDS("../output/uit_fit/fit_uit_nodiag.RDS")
mout_diag1  <- readRDS("../output/uit_fit/fit_uit1.RDS")
```

Look at the new covariance matrix. Centered a little below 1 with one weird individual. Possibley two groups of individuals.
```{r}
qplot(diag(mout_nodiag$cor_mat), bins = 30) + theme_bw()
```

Get similar correlation plots
```{r}
Rnodiag <- cov2cor(mout_nodiag$cor_mat)
corrplot(Rnodiag, method = "color", order = "hclust", diag = FALSE, title = "Not constraining diagonal", tl.pos = "n")
corrplot(mout_diag1$cor_mat, method = "color", order = "hclust", diag = FALSE, title = "Constraining diagonal", tl.pos = "n")
```

The clusters are still apparent, but less so in the Genetic correlation plot using the MAP dosage.
```{r}
genocor <- cor(t(mout_nodiag$map_dosage))
corrplot(genocor, method = "color", order = "hclust", tl.pos = "n", title = "Correlation from MAP dosage")
qplot(genocor[upper.tri(genocor)], bins = 30) + theme_bw()
```

```{r}
qplot(Rnodiag[upper.tri(Rnodiag)], bins = 30) + theme_bw()
```


# PCA on mean dosage
```{r}
library(FactoMineR)
sout <- summary(mout_nodiag)
pout <- PCA(sout$mean_dosage[-4, ])
```


```{r}
sessionInfo()
```

