---
title: "uitdewilligen"
author: "David Gerard"
date: "December 20, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

I try out mupdog on the data from

Uitdewilligen, Jan GAML, et al. "[A next-generation sequencing method for genotyping-by-sequencing of highly heterozygous autotetraploid potato.](http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0062355)" PLoS One 8.5 (2013): e62355.

Which may be obtained here:

https://doi.org/10.1371/journal.pone.0062355

Here, I assume the individuals are diploid. I want to compare the objective functions.

# Run mupdog
```{r}
library(mupdog)
library(updog)
library(tidyverse)
```

```{r, eval = FALSE}
inlist <- readRDS("../data/uitdewilligen_snps.RDS")
refmat  <- inlist$refmat
sizemat <- inlist$sizemat
ploidy  <- 2 ## major difference here.
```


```{r, eval = FALSE}
mout <- mupdog(refmat = refmat, sizemat = sizemat, ploidy = ploidy, verbose = TRUE, num_clust = 7,
               control = list(obj_tol = 10^-5, itermax = 200))
saveRDS(mout, "../output/uit_fit/fit_uit_2ploid.RDS")
```

# Analyze Output

```{r}
mout2 <- readRDS("../output/uit_fit/fit_uit_2ploid.RDS")
mout4 <- readRDS("../output/uit_fit/fit_uit1.RDS")
```


The tetraploid objective is much higher. So this is good.
```{r}
mout2$obj
mout4$obj
```


Look at a few random diploid fits
```{r}
set.seed(1)
for (index in 1:10){
  pl <- plot(mout2, sample(x = 1:length(mout2$bias), size = 1))
  print(pl)
}
```


```{r}
sessionInfo()
```



